---
title: "HTML in GT table - mtcars"
author: "Andrew vanderWilden"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r}
library(kableExtra)
library(gt)
library(sparkline)
library(glue)
library(espnscrapeR)
library(tidyverse)
```


```{r}
df <- mtcars
```


```{r}
mpg_list <- split(df$mpg, df$cyl)

mpg_list
```

```{r}
# KableExtra for boxplot in table
data.frame(
  cyl = c(4,6,8),
  boxplot = ''
) %>%
  kbl(booktabs = TRUE) %>%
  kable_paper(full_width = FALSE) %>%
  column_spec(2, image = spec_boxplot(mpg_list, width = 300, height = 70))
```


### NOTE: THIS ONLY WORKS WITH EXTERNAL OBJECT, NOT INFO IN DATAFRAME ITSELF

```{r}
# gt and KableExtra for sparklines

mpg_range <- range(df$mpg)

df %>%
  group_by(cyl) %>%
  summarise(data = list(mpg), .groups = 'drop') %>%
  mutate(
    # spec_plot creates plot in base R and returns as svg or pdf
    plot = map(data, ~spec_plot(.x, ylim = mpg_range,
                                same_lim = TRUE, width = 300, height = 70)),
    plot = map(plot, 'svg_text'),
    # need to let gt know to treat svg_text as html
    plot = map(plot, gt::html)
  ) %>%
  select(-data) %>%
  gt()
```


```{r}
# Do it all in gt

# custom function
gt_plot <- function(table_data, column, plot_data, plot_fun, ...){
  text_transform(
    table_data,
    # note the use of {{}} here - this is tidy eval
    # that allows you to indicate specific columns
    locations = cells_body(columns = vars({{column}})),
    fn = function(x){
      plot <- map(plot_data, plot_fun, width = 300, height = 70, same_lim = T, ...)
      plot_svg <- map(plot, 'svg_text')
      map(plot_svg, gt::html)
    }
  )
}


tibble(cyl = c(4,6,8), boxplot = '') %>%
  gt() %>%
  gt_plot(
    column = boxplot, # column to create boxplot in
    plot_data = mpg_list, # external data to reference
    plot_fun = spec_boxplot, # which plot function
    lim = mpg_range # range applied
  )
```



```{r}
# can quickly change to sparkline instead of boxplot

tibble(cyl = c(4,6,8), sparkline = '') %>%
  gt() %>%
  gt_plot(
    column = sparkline, # column to create plot in
    plot_data = mpg_list, # external data to reference
    plot_fun = spec_plot, # which plot function
    lim = mpg_range, # range applied
    col = 'black', # color of sparkline
    cex = 8 # change size of points
  )
```


```{r}
# put all of them together

#Note that we are just varying the ylim on the line/points 1 vs 2, where the
#mpg_line1/mpg_points1 share a common y-axis, and line2/points2 have their own y-axis.

tibble(
  cyl = c(4,6,8),
  boxplot = '',
  mpg_hist = '',
  mpg_line1 = '',
  mpg_line2 = '',
  mpg_points1 = '',
  mpg_points2 = '',
  mpg_poly = ''
) %>%
  gt() %>%
  gt_plot(column = boxplot, plot_data = mpg_list,
          plot_fun = spec_boxplot, lim = mpg_range) %>%
  gt_plot(column = mpg_hist, plot_data = mpg_list,
          plot_fun = spec_hist, lim = mpg_range) %>%
  gt_plot(column = mpg_line1, plot_data = mpg_list,
          plot_fun = spec_plot, ylim = mpg_range) %>%
  gt_plot(column = mpg_line2, plot_data = mpg_list,
          plot_fun = spec_plot) %>%
  gt_plot(column = mpg_points1, plot_data = mpg_list,
          plot_fun = spec_plot, type = 'p', ylim = mpg_range, cex = 4) %>%
  gt_plot(column = mpg_points2, plot_data = mpg_list,
          plot_fun = spec_plot, type = 'p', cex = 4) %>%
  gt_plot(column = mpg_poly, plot_data = mpg_list,
          plot_fun = spec_plot, polymin = 5, ylim = mpg_range)

```


### Use purrr::pluck() to reference actual dataframe

```{r}

df %>%
  group_by(cyl) %>%
  summarise(mpg_data = list(mpg), .groups = 'drop') %>%
  gt() %>%
  text_transform(
    locations = cells_body(columns = vars(mpg_data)),
    fn = function(x){
      # _data is data in gt table, see str() of gt object
      data_in = pluck(., '_data', 'mpg_data')
      plot = map(data_in, ~spec_plot(.x, ylim = mpg_range, same_lim = T,
                                     width = 300, height = 70))
      plot = map_chr(plot, 'svg_text')
    }
  )

```



```{r}
# alternative without going into gt object

df %>%
  group_by(cyl) %>%
  summarise(mpg_data = list(as.double(mpg)), .groups = 'drop') %>%
  gt() %>%
  text_transform(
    locations = cells_body(columns = vars(mpg_data)),
    fn = function(x){
      # split the strings at each comma
      split_data <- str_split(x, ', ')
      # convert to type double
      data <- map(split_data, as.double)
      # create the plot
      plot <- map(data, ~spec_plot(.x, ylim = mpg_range, same_lim = T,
                                   width = 300, height = 70))
      # extract the svg item
      map(plot, 'svg_text')
    }
  )

```


### Rewrite gt_plot() function so it works with a dataframe

```{r}

gt_plot <- function(table_data, plot_col, data_col, plot_fun, ...){
  # save the data extract ahead of time
  # to be used in our anonymous function below
  data_in = pluck(table_data, '_data', data_col)
  
  text_transform(
    table_data,
    # not the use of {{}} here - this is tidy eval
    # that allow you to indicate specific columns
    locations = cells_body(columns = vars({{plot_col}})),
    fn = function(x){
      plot <- map(data_in, plot_fun, width = 300, height = 70, same_lim = F, ...)
      plot_svg <- map(plot, 'svg_text')
      map(plot_svg, gt::html)
    }
  )
}
```

### Now it works with grouped list data columns

```{r}
df %>%
  group_by(cyl) %>%
  summarise(mpg_data = list(mpg), .groups = 'drop') %>%
  gt() %>%
  # note you can leave mpg_data unquoted for tidy eval
  # but have to quote mpg_data for the pluck
  gt_plot(mpg_data, 'mpg_data', plot_fun = spec_plot)

```



```{r, include=FALSE}
sparkline::sparkline(0)
```

### Interactive sparkline with javascript

```{r}

tibble(
  var = c("mpg", "wt"),
  sparkline1 = "",
  sparkline2 = "",
  box = ""
) %>% 
  gt() %>% 
  text_transform(
    locations = cells_body(vars(sparkline1)),
    fn = function(x){
      sparkline <- map(list(mtcars$mpg, mtcars$wt), ~spk_chr(values = .x, chartRangeMin = 0))
      map(sparkline, gt::html)
    }
  ) %>% 
  text_transform(
    locations = cells_body(vars(sparkline2)),
    fn = function(x){
      sparkline <- map(list(mtcars$mpg, mtcars$wt), ~spk_chr(values = .x, type = "bar", chartRangeMin = 0))
      map(sparkline, gt::html)
    }
  ) %>% 
  text_transform(
    locations = cells_body(vars(box)),
    fn = function(x){
      sparkline <- map(list(mtcars$mpg, mtcars$wt), ~spk_chr(values = .x, type = "box", chartRangeMin = 0))
      map(sparkline, gt::html)
    }
  )
```


### Sparkline function for gt()

```{r}
gt_spark <- function(table_data, plot_col, data_col){
  # save the data extract ahead of time 
  # to be used in our anonymous function below
  data_in = pluck(table_data, "_data", data_col)
  
  text_transform(
    table_data,
    # note use of {{}}, same as above for tidy eval
    locations = cells_body(columns = vars({{plot_col}})),
    fn = function(x){
      sparkline_plot <- map(
        data_in,
        ~spk_chr(values = .x, chartRangeMin = 0)
      )
      
      map(sparkline_plot, gt::html)
    }
  )
}
```


```{r}
# Apply sparkline function in gt table

df %>%
  group_by(cyl) %>%
  summarise(mpg_data = list(mpg), .groups = 'drop') %>%
  gt() %>%
  gt_spark(mpg_data, 'mpg_data')
```


### gt Example with simulated data for medical trial

```{r, echo = FALSE}
coef_table <- tibble(
  group = c(
    "",
    rep("Sex", 2),
    rep("Age", 4),
    rep("Body-Mass index", 2),
    rep("Race", 3),
    rep("Baseline statin treatment", 2),
    rep("Intensity of statin treatment", 2),
    rep("Metabolic disease", 3),
    rep("Renal function", 3)
  ),
  subgroup = c(
    "All Patients",
    "Male", "Female",
    "<65 yr", ">= 65 yr", "<75 yr", ">=75 yr",
    "<=Median", ">Median",
    "White", "Black", "Other",
    "Yes", "No",
    "High", "Not high",
    "Diabetes", "Metabolic syndrome", "Neither",
    "Normal", "Mild impairment", "Moderate impairment"
  ),
  Inclisiran = c(
    781, 535,246,297,484,638,143,394,387,653,110,18,701,80,538,243,371,195,215,395,269,113
  ),
  Placebo = c(
    780,548,232,333,447,649,131,385,394,685,87,8,692,88,546,234,331,207,242,410,260,107
  ),
  coefficients = c(-60,-55,-68,-58,-55,-57,-58,-55,-48,-58,-57,-49,-44,-58,-55,-57,-54,-52,-54,-53, -54,-52)
  ) %>% 
  mutate(
    conf_range = runif(22, min = 5, max = 10),
    conf_lower = coefficients - conf_range,
    conf_higher = coefficients + conf_range
  ) %>%
  mutate(
    image = spec_pointrange(
      x = coefficients, 
      xmin = conf_lower, 
      xmax = conf_higher, 
      same_lim = TRUE,
      lim = c(-100, 25),
      vline = 0,
      width = 550,
      cex = .75,
      col = "black"
      )
    )
```

```{r}
coef_table %>%
  select(-coefficients, -contains('conf')) %>%
  mutate(
    image = map(image, 'svg_text'),
    image = map(image, ~gt::html(as.character(.x)))
  ) %>%
  select(group:Placebo, pct_diff = image) %>%
  gt(
    groupname_col = 'group',
    rowname_col = 'subgroup'
  ) %>%
  opt_row_striping() %>%
  tab_options(
    data_row.padding = px(3)
  )
```


```{r, echo = FALSE}
# Add tooltip to column labels
library(htmltools)

with_tooltip <- function(value, tooltip) {
  tags$abbr(style = "text-decoration: underline; text-decoration-style: solid; cursor: question; color: blue",
            title = tooltip, value)
}
```



```{r, echo = FALSE}
# creates a star rating 0-5

rating_stars <- function(rating, max_rating = 5) {
  rounded_rating <- floor(rating + 0.5)  # always round up
  stars <- lapply(seq_len(max_rating), function(i) {
    if (i <= rounded_rating) fontawesome::fa("star", fill= "orange") else fontawesome::fa("star", fill= "grey")
  })
  label <- sprintf("%s out of %s", rating, max_rating)
  div_out <- div(title = label, "aria-label" = label, role = "img", stars)
  
  as.character(div_out) %>% 
    gt::html()
}
```



```{r,echo=FALSE}
# fontawesome for inline icons

rank_chg <- function(change_dir){
  if (change_dir == "increase") {
    logo_out <- fontawesome::fa("arrow-up", fill = "blue")
  } else if (change_dir == "decrease"){
    logo_out <- fontawesome::fa("arrow-down", fill = "red")
  }
  
  logo_out %>% 
    as.character() %>% 
    gt::html()
  
}
```



```{r, echo = FALSE}
# create badge style label with rounded edges and specific color

add_cyl_color <- function(cyl){
      add_color <- if (cyl == 4) {
        "background: hsl(116, 60%, 90%); color: hsl(116, 30%, 25%);"
      } else if (cyl == 6) {
        "background: hsl(230, 70%, 90%); color: hsl(230, 45%, 30%);"
      } else if (cyl == 8) {
        "background: hsl(350, 70%, 90%); color: hsl(350, 45%, 30%);"
      }
      div_out <- htmltools::div(
        style = paste(
          "display: inline-block; padding: 2px 12px; border-radius: 15px; font-weight: 600; font-size: 12px;",
          add_color
          ),
        paste(cyl, "Cylinders")
      )
      
      as.character(div_out) %>% 
        gt::html()
}
```


```{r, echo = FALSE}
# inline barchart

bar_chart <- function(value, color = "red"){
    
    glue::glue("<span style=\"display: inline-block; direction: ltr; border-radius: 4px; padding-right: 2px; background-color: {color}; color: {color}; width: {value}%\"> &nbsp; </span>") %>% 
    as.character() %>% 
    gt::html()
}
```






### Final mtcars table

```{r}
set.seed(377)

df %>%
  tibble() %>%
  select(1:4) %>%
  sample_n(size = 6) %>%
  mutate(
    rank_change = sample(c('increase', 'decrease'), size = 6, replace = T),
    rank_change = map(rank_change, rank_chg)
  ) %>%
  mutate(
    rating = sample(1:5, size = 6, replace = T),
    rating = map(rating, rating_stars)
  ) %>%
  mutate(
    cylinder = map(cyl, add_cyl_color)
  ) %>%
  mutate(
    mpg_plot = mpg/max(mpg) * 100,
    mpg_plot = map(mpg_plot, ~bar_chart(value = .x, color = 'lightblue'))
  ) %>%
  gt() %>%
  cols_align(
    align = 'left',
    columns = vars(mpg_plot)
  ) %>%
  cols_label(
    mpg = gt::html(as.character(with_tooltip('MPG', 'Miles per Gallon')))
  ) %>%
  tab_source_note(
    source_note = html(
      htmltools::tags$a(
        href = 'https://gt.rstudio.com/reference/md.html',
        target = '_blank',
        'Data Source'
      ) %>%
        as.character()
    )
  ) %>%
  tab_source_note(
    source_note = html(
      "<details><h3 style='font-face:bold'>Table Key</h3><div>MPG: Miles Per Gallon</div><div>Cyl: Cylinders</div><div>disp: Displacement</div><div>hp: Horsepower</div><div>rank_change: Rank Change</div><div>rating: Rating</div></details>"

    )
  ) %>%
  tab_options(
    data_row.padding = px(5)
  )
```



# Second Example with NFL QBs
  
```{r}
# use espnscrapeR to get NFL standings + QBR ratings
nfl_qbr <- get_nfl_qbr(2020)
nfl_standings <- get_nfl_standings(2020)

# also get weekly for embedded plot
qbr_weekly <- crossing(season = 2020, week = 1:8) %>%
  pmap_dfr(.f = get_nfl_qbr)
```


```{r}
# Data prep

qbr_match <- qbr_weekly %>%
  filter(short_name %in% nfl_qbr$short_name) %>%
  group_by(short_name, team) %>%
  summarise(qbr_weekly = list(qbr_total), .groups = "drop",
            qbr = mean(qbr_total),
            qbr_sd = sd(qbr_total),
            plays = sum(qb_plays),
            pass = mean(pass),
            run = mean(run),
            head = unique(headshot_href),
            n = n()) %>%
  arrange(desc(qbr)) %>% 
  filter(n >= 7)

# clean up the data a bit and combine
tab_df <- qbr_match %>% 
  left_join(nfl_standings, by = c("team" = "abb_name")) %>%
  select(short_name, team, head, qbr_weekly:run, wins, losses, points_for) %>%
  mutate(wl = glue("{wins}-{losses}")) %>%
  select(-wins, -losses)
tab_df
```

```{r}
# calc rank change
qbr_rnk_chg <- qbr_weekly %>% 
  mutate(game_week = as.integer(game_week)) %>% 
  group_by(short_name) %>% 
  mutate(mean_qbr = mean(qbr_total)) %>% 
  ungroup() %>% 
  select(game_week, rank, short_name, qbr_total, mean_qbr) %>% 
  filter(game_week != max(game_week)) %>% 
  filter(short_name %in% nfl_qbr$short_name) %>%
  group_by(short_name) %>%
  summarize(prev_qbr = mean(qbr_total), mean_qbr = unique(mean_qbr)) %>% 
  mutate(
    prev_week = rank(-prev_qbr),
    rank = rank(-mean_qbr)
    ) %>% 
  mutate(rank_chg = prev_week-rank) %>% 
  ungroup() %>% 
  arrange(desc(mean_qbr)) %>% 
  select(short_name, qbr = mean_qbr, rank_chg, rank)

qbr_rnk_chg

```

```{r}
# Code for name, team, record combo

combine_word <- function(name, team, wl){
      glue::glue(
        "<div style='line-height:10px'><span style='font-weight:bold;font-variant:small-caps;font-size:14px'>{name}</div>
        <div style='line-height:12px'><span style ='font-weight:bold;color:grey;font-size:10px'>{team}&nbsp;&nbsp;{wl}</span></div>"
      )
    }

combo_df <- tab_df %>% 
  left_join(qbr_rnk_chg, by = c("short_name", "qbr")) %>%
  select(rank, rank_chg, short_name:wl) %>% 
  mutate(
    rank = row_number(),
    combo = combine_word(short_name, team, wl),
    combo = map(combo, gt::html)
    ) %>% 
  select(rank, rank_chg, head, combo, qbr, qbr_weekly, plays, points_for)



```


```{r}
final_table <- combo_df %>%
  gt() %>%
  cols_align(
    align = 'left',
    columns = vars(combo)
  ) %>%
  tab_options(
    data_row.padding = px(2)
  ) %>%
  text_transform(
    locations = cells_body(columns = vars(head)),
    fn = function(x){
      gt::web_image(x)
    }
  ) %>%
    text_transform(
    locations = cells_body(columns = vars(rank_chg)),
    fn = function(x){
      
      rank_chg <- as.integer(x)
      
      choose_logo <-function(x){
        if (x == 0){
        gt::html(fontawesome::fa("equals", fill = "grey"))
      } else if (x > 0){
         gt::html(glue::glue("<span style='color:#1134A6;font-face:bold;font-size:10px;'>{x}</span>"), fontawesome::fa("arrow-up", fill = "#1134A6"))
      } else if (x < 0) {
        gt::html(glue::glue("<span style='color:#DA2A2A;font-face:bold;font-size:10px;'>{x}</span>"), fontawesome::fa("arrow-down", fill = "#DA2A2A"))
      }
      } 
      
      map(rank_chg, choose_logo)
    
    }
  ) %>% 
  fmt_number(
    columns = vars(qbr), 
    decimals = 1
    ) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(TRUE)
  ) %>% 
  cols_label(
    rank = "RK",
    combo = "",
    head = "QB",
    qbr = "QBR",
    plays = "PLAYS",
    points_for = "PF",
    qbr_weekly = "WEEKLY",
    rank_chg = ""
  ) %>% 
  gt_spark(qbr_weekly, "qbr_weekly") %>%
  espnscrapeR::gt_theme_espn() %>% 
  tab_source_note(
    source_note = gt::html(
      htmltools::tags$a(
        href = "https://www.espn.com/nfl/qbr", 
        target = "_blank", 
        "Data: ESPN"
        ) %>% 
        as.character()
      )
    ) %>% 
  cols_align(
    "left",
    columns = vars(qbr_weekly)
  ) %>% 
  cols_width(
    vars(rank) ~ px(25),
    vars(rank_chg) ~ px(35),
    vars(head) ~ px(50),
    vars(combo) ~ px(115),
    vars(qbr) ~ px(35),
    vars(plays) ~ px(50),
    vars(points_for) ~ px(35),
    vars(qbr_weekly) ~ px(75)
  ) %>% 
  tab_header(
    title = gt::html("<h3>NFL QBR through Week 8</h3>")
  ) %>% 
  tab_options(
    table.width = px(480),
    data_row.padding = px(4)
  )
final_table
```



